Faithful Variable Screening for\\ High-Dimensional Convex Regression

### 2014-09-29 tex/scam/additive_faithful.tex
boundary flatness name change

\begin{definition}
\label{defn:boundary-point}
A density $p(\mathbf{x})$ be a density supported on $[0,1]^p$ satisfies
the \emph{boundary points condition} if, for all $j$, and for all $\mathbf{x}_{-j}$:

@@

The boundary points condition is a weak condition. For instance, it is
satisfied when the density is flat at the boundary of support---more
precisely, when the joint density satisfies the property that
$\frac{\partial p(x_j,\mathbf{x}_{-j})}{\partial x_j} =
\frac{\partial^2 p(x_j, \mathbf{x}_{-j})}{\partial x_j^2} = 0$ at
points $x_j = 0, x_j=1$. The boundary points property is also
trivially satisfied when $p$ is a product density.

@@

\begin{theorem}
\label{thm:convex_faithful}
Let $p$ be a positive density supported on $C=[0,1]^p$ that satisfies
the boundary points property. 
If $f$ is convex and twice differentiable, then $f$ is additively faithful under $p$.
\end{theorem}

@@

By the boundary points condition, we have that $p''(\mathbf{x}_{-k}
\given x_k)$ and $p'(\mathbf{x}_{-k} \given x_k)$ are zero at $x_k =
x_k^0 \equiv 0$. The integral equations then reduce to the following:
\begin{align}
& \int_{\mathbf{x}_{-k}} p(\mathbf{x}_{-k} \given x^0_k) f'(x^0_k, \mathbf{x}_{-k}) d \mathbf{x}_{-k}= 0 \label{eqn:integral1b} \\
& \int_{\mathbf{x}_{-k}} p(\mathbf{x}_{-k} \given x^0_k) f''(x^0_k, \mathbf{x}_{-k}) d\mathbf{x}_{-k} = 0.
\end{align}
Because $f$ is convex, $f(x_k, \mathbf{x}_{-k})$ must be a convex
function of 
$x_k$ for all $\mathbf{x}_{-k}$. Therefore, for all $\mathbf{x}_{-k}$,
$f''(x^0_k, \mathbf{x}_{-k}) \geq 0$. Since $p(\mathbf{x}_{-k} \given
x^0_k) > 0$ by the assumption that $p$ is a positive density, 
we have that $\forall \mathbf{x}_{-k}, f''(x^0_k, \mathbf{x}_{-k}) = 0$ necessarily.

@@

Interestingly, it is possible to approximate the Gaussian distribution arbitrarily well with distributions that do satisfy the boundary-points conditions.

\begin{example} 
\label{examp:boundaryflat_example}
 Consider a mixture $\lambda U[-(b+\epsilon), b+\epsilon]^2 + (1-\lambda) N_b(0, \Sigma)$ where $N_b(0,\Sigma)$ is the density of a \emph{truncated} bivariate Gaussian bounded in $[-b, b]^2$. The uniform distribution is supported over a slightly larger square to satisfy the boundary-points conditions.

@@

\begin{figure*}[htp]
\vskip-10pt
	\centering
	\subfigure[Gaussian distribution]{
          \label{fig:gaussian_example}
		\centering
		{\includegraphics[width=0.4\textwidth]{figs/gaussian_example}}
	}
	\subfigure[Mixture approximation]{
          \label{fig:boundaryflat_example}
		\centering
		{\includegraphics[width=0.4\textwidth]{figs/boundaryflat_example}}
	}
\caption{Optimal additive projection of the quadratic function described in Example~\ref{examp:gaussian_counterexample} under both the Gaussian distribution described in Example~\ref{examp:gaussian_counterexample} and under the approximately Gaussian mixture distribution described in Example~\ref{examp:boundaryflat_example}. For the mixture approximation, we used $b=5, \epsilon=0.3, \lambda=0.0001$ where the parameters are defined in Example~\ref{examp:boundaryflat_example}. This example shows the effect and the importance of the boundary-points conditions.}
\vskip-10pt
\end{figure*}

@@

\begin{theorem}
\label{thm:acdc_faithful}
Suppose $p(\mathbf{x})$ is a positive density on $C=[0,1]^p$ that
satisfies the boundary points condition. Suppose that $f$ is convex
and twice differentiable, and that $\partial_{x_k} p( \mathbf{x}_{-k} \given x_k )$, and
$\partial_{x_k}^2 p( \mathbf{x}_{-k} \given x_k)$ are all continuous
as functions of $x_k$.  Let $f^*_k$ and $g^*_k$ be as defined in equations
\eqref{eqn:convex_additive} and \eqref{eqn:concave_postprocess}.  Then
$f^*_k = 0$ and $g^*_k = 0$ implies that $\partial_{x_k} f(\mathbf{x})
= 0$, that is, $f$ does not depend on~$x_k$.
\end{theorem}

\begin{lemma}
\label{cor:faithfulness_extension}
Suppose $p(x)$ is a positive density on $C=[0,1]^p$ satisfying the boundary points condition.
Let $f(\mathbf{x})$ be a convex twice differentiable function on $C$. Let $\phi(\mathbf{x}_{-k})$ be any function that does not depend on $x_k$. Then, we have that the unconstrained univariate function
 \begin{align}
h^*_k = \arg\min_{f_k} \mathbb{E} \Bigl[\bigl( f(X) 
           - \phi(X_{-k}) - h_k(X_k) \bigr)^2\Bigr]
\end{align}
is given by $h^*_k(x_k) = \mathbb{E}\bigl[ f(X) - \phi(X_{-k}) \,|\, x_k\bigr]$,
and $ h^*_k = 0$ implies that $\partial_{x_k} f(x) = 0$.
\end{lemma}

>>>>

\begin{definition}
\label{defn:boundary-point}
A density $p(\mathbf{x})$ be a density supported on $[0,1]^p$ satisfies
the \emph{boundary flatness condition} if, for all $j$, and for all $\mathbf{x}_{-j}$:

@@

The boundary flatness condition is a weak condition. For instance, it is
satisfied when the density is flat at the boundary of support---more
precisely, when the joint density satisfies the property that
$\frac{\partial p(x_j,\mathbf{x}_{-j})}{\partial x_j} =
\frac{\partial^2 p(x_j, \mathbf{x}_{-j})}{\partial x_j^2} = 0$ at
points $x_j = 0, x_j=1$. The boundary flatness property is also
trivially satisfied when $p$ is a product density.

@@

\begin{theorem}
\label{thm:convex_faithful}
Let $p$ be a positive density supported on $C=[0,1]^p$ that satisfies
the boundary flatness property. 
If $f$ is convex and twice differentiable, then $f$ is additively faithful under $p$.
\end{theorem}

@@

By the boundary flatness condition, we have that $p''(\mathbf{x}_{-k}
\given x_k)$ and $p'(\mathbf{x}_{-k} \given x_k)$ are zero at $x_k =
x_k^0 \equiv 0$. The integral equations then reduce to the following:
\begin{align}
& \int_{\mathbf{x}_{-k}} p(\mathbf{x}_{-k} \given x^0_k) f'(x^0_k, \mathbf{x}_{-k}) d \mathbf{x}_{-k}= 0 \label{eqn:integral1b} \\
& \int_{\mathbf{x}_{-k}} p(\mathbf{x}_{-k} \given x^0_k) f''(x^0_k, \mathbf{x}_{-k}) d\mathbf{x}_{-k} = 0.
\end{align}
Because $f$ is convex, $f(x_k, \mathbf{x}_{-k})$ must be a convex
function of 
$x_k$ for all $\mathbf{x}_{-k}$. Therefore, for all $\mathbf{x}_{-k}$,
$f''(x^0_k, \mathbf{x}_{-k}) \geq 0$. Since $p(\mathbf{x}_{-k} \given
x^0_k) > 0$ by the assumption that $p$ is a positive density, 
we have that $\forall \mathbf{x}_{-k}, f''(x^0_k, \mathbf{x}_{-k}) = 0$ necessarily.

@@

Interestingly, it is possible to approximate the Gaussian distribution arbitrarily well with distributions that do satisfy the boundary flatness conditions.

\begin{example} 
\label{examp:boundaryflat_example}
 Consider a mixture $\lambda U[-(b+\epsilon), b+\epsilon]^2 + (1-\lambda) N_b(0, \Sigma)$ where $N_b(0,\Sigma)$ is the density of a \emph{truncated} bivariate Gaussian bounded in $[-b, b]^2$. The uniform distribution is supported over a slightly larger square to satisfy the boundary flatness conditions.

@@

\begin{figure*}[htp]
\vskip-10pt
	\centering
	\subfigure[Gaussian distribution]{
          \label{fig:gaussian_example}
		\centering
		{\includegraphics[width=0.4\textwidth]{figs/gaussian_example}}
	}
	\subfigure[Mixture approximation]{
          \label{fig:boundaryflat_example}
		\centering
		{\includegraphics[width=0.4\textwidth]{figs/boundaryflat_example}}
	}
\caption{Optimal additive projection of the quadratic function described in Example~\ref{examp:gaussian_counterexample} under both the Gaussian distribution described in Example~\ref{examp:gaussian_counterexample} and under the approximately Gaussian mixture distribution described in Example~\ref{examp:boundaryflat_example}. For the mixture approximation, we used $b=5, \epsilon=0.3, \lambda=0.0001$ where the parameters are defined in Example~\ref{examp:boundaryflat_example}. This example shows the effect and the importance of the boundary flatness conditions.}
\vskip-10pt
\end{figure*}

@@

\begin{theorem}
\label{thm:acdc_faithful}
Suppose $p(\mathbf{x})$ is a positive density on $C=[0,1]^p$ that
satisfies the boundary flatness condition. Suppose that $f$ is convex
and twice differentiable, and that $\partial_{x_k} p( \mathbf{x}_{-k} \given x_k )$, and
$\partial_{x_k}^2 p( \mathbf{x}_{-k} \given x_k)$ are all continuous
as functions of $x_k$.  Let $f^*_k$ and $g^*_k$ be as defined in equations
\eqref{eqn:convex_additive} and \eqref{eqn:concave_postprocess}.  Then
$f^*_k = 0$ and $g^*_k = 0$ implies that $\partial_{x_k} f(\mathbf{x})
= 0$, that is, $f$ does not depend on~$x_k$.
\end{theorem}

\begin{lemma}
\label{cor:faithfulness_extension}
Suppose $p(x)$ is a positive density on $C=[0,1]^p$ satisfying the boundary flatness condition.
Let $f(\mathbf{x})$ be a convex twice differentiable function on $C$. Let $\phi(\mathbf{x}_{-k})$ be any function that does not depend on $x_k$. Then, we have that the unconstrained univariate function
 \begin{align}
h^*_k = \arg\min_{f_k} \mathbb{E} \Bigl[\bigl( f(X) 
           - \phi(X_{-k}) - h_k(X_k) \bigr)^2\Bigr]
\end{align}
is given by $h^*_k(x_k) = \mathbb{E}\bigl[ f(X) - \phi(X_{-k}) \,|\, x_k\bigr]$,
and $ h^*_k = 0$ implies that $\partial_{x_k} f(x) = 0$.
\end{lemma}

### 2014-09-29 tex/scam/summary.tex
added comment on sufficiently large n
boundary flatness name change

Our first result shows that convex multivariate functions are
additively faithful
under the following assumption on the distribution of the data.
\begin{definition}
  Let $p(\mathbf{x})$ be a density supported on $[0,1]^p$.  Then $p$
  satisfies the \emph{boundary points condition} if for all $j$, and
  for all $\mathbf{x}_{-j}$,
\[
\frac{\partial p(\mathbf{x}_{-j} \given x_j)}{\partial x_j}  =  
\frac{\partial^2 p(\mathbf{x}_{-j} \given x_j)}{\partial x_j^2} = 0
\quad \trm{at $x_j = 0$ and $x_j = 1$}.
\]
\end{definition}

@@

\begin{stheorem}
  Let $p$ be a positive density supported on $C=[0,1]^p$ that
  satisfies the boundary points property. If $f$ is convex and twice
  differentiable, then $f$ is additively faithful under $p$.
\end{stheorem}

@@

\begin{stheorem}
Suppose $p(\mathbf{x})$ is a positive
density on $C=[0,1]^p$ that satisfies the boundary points
condition. Suppose that $f$ is convex and twice-differentiable.
and that $\partial_{x_k} f$, $\partial_{x_k} p( \mathbf{x}_{-k}
\given x_k )$, and $\partial_{x_k}^2 p( \mathbf{x}_{-k} \given x_k)$
are all continuous as functions on $C$.
Define
\begin{equation}
\{ f^*_k \}_{k=1}^p,\mu^* = \arg\min_{\{f_k\},\mu} \Big \{
\E\Bigl( f(X) - \mu - \sum_{k=1}^s f_k(X_k) \Bigr)^2 \,:\, f_k \in
\mathcal{C}^1, \, \E f_k(X_k) = 0 \Big \}
\end{equation} where $\mathcal{C}^1$ is the set of univariate convex
functions, and
\begin{equation}
g^*_k = \arg\min_{g_k} \Big\{ \E\Bigl( f(X) - \mu^* - 
\sum_{k' \neq k} f^*_{k'}(X_{k'}) - g_k \Bigr)^2 \,:\, g_k \in \mh
\mathcal{C}^1, \E g_k(X_k) = 0 \Big\},
\end{equation}
with $\mh{}\mathcal{C}^1$ denoting the set of univariate concave
functions.  Then $f^*_k = 0$ and $g^*_k = 0$ implies that $f$ does not
depend on $x_k$, i.e., $\partial_{x_k} f(\mathbf{x}) = 0$ with
probability one.
\end{stheorem}

@@

>>>>

Our first result shows that convex multivariate functions are
additively faithful
under the following assumption on the distribution of the data.
\begin{definition}
  Let $p(\mathbf{x})$ be a density supported on $[0,1]^p$.  Then $p$
  satisfies the \emph{boundary flatness condition} if for all $j$, and
  for all $\mathbf{x}_{-j}$,
\[
\frac{\partial p(\mathbf{x}_{-j} \given x_j)}{\partial x_j}  =  
\frac{\partial^2 p(\mathbf{x}_{-j} \given x_j)}{\partial x_j^2} = 0
\quad \trm{at $x_j = 0$ and $x_j = 1$}.
\]
\end{definition}

@@

\begin{stheorem}
  Let $p$ be a positive density supported on $C=[0,1]^p$ that
  satisfies the boundary flatness property. If $f$ is convex and twice
  differentiable, then $f$ is additively faithful under $p$.
\end{stheorem}

@@

\begin{stheorem}
Suppose $p(\mathbf{x})$ is a positive
density on $C=[0,1]^p$ that satisfies the boundary flatness
condition. Suppose that $f$ is convex and twice-differentiable.
and that $\partial_{x_k} f$, $\partial_{x_k} p( \mathbf{x}_{-k}
\given x_k )$, and $\partial_{x_k}^2 p( \mathbf{x}_{-k} \given x_k)$
are all continuous as functions on $C$.
Define
\begin{equation}
\{ f^*_k \}_{k=1}^p,\mu^* = \arg\min_{\{f_k\},\mu} \Big \{
\E\Bigl( f(X) - \mu - \sum_{k=1}^s f_k(X_k) \Bigr)^2 \,:\, f_k \in
\mathcal{C}^1, \, \E f_k(X_k) = 0 \Big \}
\end{equation} where $\mathcal{C}^1$ is the set of univariate convex
functions, and
\begin{equation}
g^*_k = \arg\min_{g_k} \Big\{ \E\Bigl( f(X) - \mu^* - 
\sum_{k' \neq k} f^*_{k'}(X_{k'}) - g_k \Bigr)^2 \,:\, g_k \in \mh
\mathcal{C}^1, \E g_k(X_k) = 0 \Big\},
\end{equation}
with $\mh{}\mathcal{C}^1$ denoting the set of univariate concave
functions.  Then $f^*_k = 0$ and $g^*_k = 0$ implies that $f$ does not
depend on $x_k$, i.e., $\partial_{x_k} f(\mathbf{x}) = 0$ with
probability one.
\end{stheorem}

@@

%% this statement "for sufficiently large n" should be changed.  the
%% result is stronger than this, since we get a sample complexity.
